---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4

In this assignment we do the following:
- we reproduce the meta-analysis of pitch SD from last semester in a Bayesian framework
- we reproduce the pitch SD in schizophrenia analysis from last semester using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Reproduce the meta-analysis of pitch sd from previous studies of voice in schizophrenia
- the data is available as Assignment4MetaData.xlsx
- Effect size (cohen's d), sd and variance are already calculated (you're welcome!)
- Since we're only interested in getting a meta-analytic effect size, let's take a shortcut and use bromance magic (brms): https://mvuorre.github.io/post/2016/2016-09-29-bayesian-meta-analysis/

Step 2: Prepare the pitch SD data from last year
- the data is available as Assignment4PitchData.csv (thanks Celine)
- We do not know how to build random effects, yet. So we need to simplify the dataset to avoid interdependence between datapoint: How?
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).

Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality

Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality

Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare their relative distance from truth (WAIC)
- Discuss how they compare and whether any of them is best.

Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using WAIC.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?

Optional step 9: Bromance magic.
- explore the bromance code below including random effects (by default with weakly informative priors)
- learn how to change the prior
- explore effects of trial, age, gender, including the appropriate random slopes
- compare the models you created using WAIC and posterior predictive check (pp_check())


```{r}

Sys.getenv("PATH")
 system('g++ -v')
 
 cat('Sys.setenv(BINPREF = "C:/Rtools/mingw_$(WIN)/bin/")',
    file = file.path(Sys.getenv("HOME"), ".Rprofile"), 
    sep = "\n", append = TRUE)
 install.packages("rstan", repos = "https://cloud.r-project.org/", lib = 'C:/Users/kilda/Documents/R/win-library/3.5', dependencies=TRUE)

library(rstan)
library(pacman)
p_load(tidyverse, rethinking, pastecs, patchwork, Brobdingnag, metafor, brms, lme4, ggplot)
setwd("C:/Users/kilda/OneDrive/Documents/R")
devtools::install_github("mvuorre/brmstools")
library(readxl)
library(brms)

dat = read.csv("Assignment4MetaData.csv", stringsAsFactors = F)
devtools::install_github("thomasp85/patchwork", force = TRUE)

# dataxl = read_excel("Assignment4MetaData.xlsx")
pitch_data = read.csv("Assignment4PitchDatav2.csv")

dat$MeanES = as.numeric(dat$MeanES)
dat$SdES = as.numeric(dat$SdES)
dat$StudyID = as.numeric(dat$StudyID)

library(metafor)
ma_out <- rma(data = dat, yi = MeanES, sei = SdES, slab = dat$StudyID)
summary(ma_out)

# dat$StudyRef = as.numeric(dat$StudyRef)

sub_dat = subset(dat, !is.na(MeanES))


brm_out <- brm( MeanES | se(SdES) ~ 1 + (1|StudyID),
               data = dat, iter = 2000, chain = 2, cores = 2)
brm_out
# prior = set_prior("uniform(0, 1000)", can set in, 

# plot 
plot(brm_out)

# forest plot 
# library(brmstools)
forest(brm_out)

```

QUESTION 2 
```{r}
library(dplyr)
p_sum = pitch_data %>%
  group_by(ID_unique) %>%
  summarise(ID = mean(ID), Diagnosis = mean(diagnosis), Studynr = mean(studynr), PitchMean = mean(PitchMean), PitchSD = mean(PitchSD), PitchMedian = mean(PitchMedian), PitchRange = mean(PitchRange), PitchIQR = mean(PitchIQR), PitchMad = mean(PitchMad), PitchCV = mean(PitchCV)) %>%
  mutate(MeanScale = scale(PitchMean), SDScale = scale(PitchSD)[,], MedianScale = scale(PitchMedian), RangeScale = scale(PitchRange), IQRScale = scale(PitchIQR), MadScale = scale(PitchMad), CVScale = scale(PitchCV))

p_sum
# new data frame we can use with all mean values to use later 

p_sum = as.data.frame(p_sum)

```

Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality
```{r}

# plotting the liklihood function 

rethinking::dens(p_sum$PitchSD[p_sum$Diagnosis == '1'], col= 'red')
rethinking::dens(p_sum$PitchSD[p_sum$Diagnosis == '0'], add = T)

# loooooose 
m1 = rethinking::map(
   alist(
    SDScale ~ dnorm( mu , sigma ) , 
    mu <- a + b*Diagnosis ,
    a ~ dnorm( 0 , 1 ) , 
    b ~ dnorm( 0 , 1) , # loose prior 
    sigma ~ dunif( 0 , 3 ) 
  ) ,
  data= p_sum)

pm1 = rethinking::precis(m1)
plot(pm1)

# weakly regularising 
m2 = rethinking::map(
   alist(
    SDScale ~ dnorm( mu , sigma ) , 
    mu <- a + b*Diagnosis ,
    a ~ dnorm( 0 , 1 ) , 
    b ~ dnorm( 0 , 0.5) , # weakly regularising prior
    sigma ~ dunif( 0 , 3 ) 
  ) ,
  data= p_sum)
pm2 = rethinking::precis(m2)
plot(pm2)


# 0.25 more conservative prior 
m3 = rethinking::map(
   alist(
    SDScale ~ dnorm( mu , sigma ) , 
    mu <- a + b*Diagnosis ,
    a ~ dnorm( 0 , 1 ) , 
    b ~ dnorm( 0 , 0.25) , # more conservative 
    sigma ~ dunif( 0 , 3 ) 
  ) ,
  data= p_sum)
pm3 = rethinking::precis(m3)
plot(pm3)

# most conservative 
m4 = rethinking::map(
   alist(
    SDScale ~ dnorm( mu , sigma ) , 
    mu <- a + b*Diagnosis ,
    a ~ dnorm( 0 , 1 ) , 
    b ~ dnorm( 0 , 0.1) , # more conservative 
    sigma ~ dunif( 0 , 3 ) 
  ) ,
  data= p_sum)
pm4 = rethinking::precis(m4)
plot(pm4)

# plotting thr effect of the priors 

gggplot = function(model, x = NULL , y = NULL, dataframe){ 
   #where: 
  #model is a model resulting from MAP 
  #x is the x variable as a string (can be left unspecified) 
  #y is the y variable as a string (can be left unspecified) 
  #dataframe is the dataframe from which the model is trained upon 
  #requires packages: dplyr, ggplot2, stringr and rethinking
  
  if (is.null(y) == T){ 
    #if y isn't specified - extract y 
    temp <- rethinking::flist_untag(model@formula)[[1]] 
    y <- as.character(temp[[2]])
    } 
  if (is.null(x) == T){ 
    #if x isn't specified - extract x 
      temp <- rethinking::flist_untag(model@formula)[[2]] 
      x <- gsub(" ","",gsub(".*\\*","",as.character(temp[[3]])[3])) 
  }
  
   #extracting data from the dataframe 
  xd <- dplyr::select(dataframe, x) 
  yd <- dplyr::select(dataframe, y)
  
   #calculate the prior data 
  post <- rethinking::extract.samples(model) 
  #posterior from samples 
  mu.link <- function(x) post$a + post$b*x 
  VIQ.seq <- seq( from=min(xd) , to=max(xd) , by=0.1 ) 
  mu <- sapply( VIQ.seq , mu.link ) 
  mu.mean <- apply( mu , 2 , mean ) 
  mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 ) 
  temp_list = list(VIQ.seq); names(temp_list) <- x 
  #naming the vector in the list (since the sim function requires it) 
  sim.VIQ <- sim( model , data=temp_list ) 
  VIQ.PI <- apply( sim.VIQ , 2 , PI , prob=0.89 )

  #transform the data from ggplot 
    #for making the line 
  temp_df <- data.frame(VIQ.seq = VIQ.seq, mu.mean = mu.mean) 
  #for making the HDPI 
  temp_matrix <- t(data.frame(mu.HPDI = mu.HPDI)) 
  #t() is a tranpose function which returns a matrix 
  temp_df1 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], mu.mean = mu.mean, VIQ.seq = VIQ.seq)
   #for making the PI 
  temp_matrix <- t(data.frame(VIQ.PI = VIQ.PI)) 
  temp_df2 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], mu.mean = mu.mean, VIQ.seq = VIQ.seq) 
  #for the points 
  dataframe = data.frame(xd, yd) 
  #taking the data to new column as to chose the name (maintaining old name for the names on the plot) 
  dataframe$c1 = dataframe[,1] 
  dataframe$c2 = dataframe[,2]
  
   #make the plot 
  grob = ggplot(dataframe, aes(x = c1, y = c2)) + 
    geom_ribbon(data = temp_df2, aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper), fill = "grey70", alpha = 2/3) + 
    geom_ribbon(data = temp_df1, aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper), fill = "grey83", alpha = 2/3) + 
    geom_line(data = temp_df, aes(x = VIQ.seq, y = mu.mean))+ 
    geom_point(color = "navyblue", shape = 1, size = 1.5, alpha = 2/3)+ 
    xlab(colnames(dataframe)[1]) + 
    ylab(colnames(dataframe)[2]) 
  #using the names for the axis 
  return(grob) 
  } #go go gadget fancy plot


m1plot = gggplot(model = m1, dataframe = p_sum)
m2plot = gggplot(model = m2, dataframe = p_sum)
m3plot = gggplot(model = m3, dataframe = p_sum)
m4plot = gggplot(model = m4, dataframe = p_sum)

m1plot + m2plot + m3plot + m4plot

```

Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality

```{r}
# most conservative 
m5 = rethinking::map(
   alist(
    SDScale ~ dnorm( mu , sigma ) , 
    mu <- a + b*Diagnosis ,
    a ~ dnorm( 0 , 1 ) , 
    b ~ dnorm( -.54 , 0.23) , # more conservative 
    sigma ~ dunif( 0 , 3 ) 
  ) ,
  data= p_sum)
precis(m5)
plot(precis(m5))

## we want to plot all the posteriors of the models, and compare the diferences 

```





```{r}
brm_out1 <- brm(PitchSD ~ 1 + Diagnosis  +(1|ID_unique/Study), # Outcome as a function of the predictors as in lme4. 
               data= pitch_data, # Define the data
               family=gaussian(), # Define the family. 
               iter = 5000, warmup = 2000, cores = 4)
summary(brm_out1)
plot(brm_out1)